Cell type assignments for SCPCP000015

### 2025-02-11 analyses/cell-type-ewings/exploratory_analysis/08-merged-celltypes.Rmd

- Annotations obtained by running `SingleR` with tumor cells as a reference output by [`aucell-singler-annotation.sh`](https://github.com/AlexsLemonade/OpenScPCA-analysis/blob/main/analyses/cell-type-ewings/aucell-singler-annotation.sh). 
- Consensus cell type annotations output by running [`assign-consensus-celltypes.sh` in `cell-type-consensus`](https://github.com/AlexsLemonade/OpenScPCA-analysis/blob/main/analyses/cell-type-consensus/assign-consensus-celltypes.sh). 
- AUC values as calculated by `AUCell` for a set of Ewing sarcoma specific gene sets in MSigDB output by [`run-aucell-ews-signatures.sh`](https://github.com/AlexsLemonade/OpenScPCA-analysis/blob/main/analyses/cell-type-ewings/run-aucell-ews-signatures.sh).

@@

Any cells that are unable to be labeled via consensus cell types are labeled as "Unknown" and I expect these will line up with cells labeled as tumor cells by `SingleR`.

@@

```{r}
# get columns with mean expression of custom gene sets 
mean_exp_columns <- colnames(all_info_df)[which(endsWith(colnames(all_info_df), "_mean"))]
# make sure order is defined by cell type order
all_info_df <- all_info_df |> 
  dplyr::arrange(final_lumped)

@@

# first get tally for how many libraries each cell type is present in 
total_libraries <- all_info_df |> 
  dplyr::select(final_annotation, library_id) |> 
  unique() |> # get rid of duplicate cells
  dplyr::group_by(final_annotation) |> 
  dplyr::summarize(
    total_libraries = length(library_id)) |> 
  dplyr::arrange(desc(total_libraries))

# print out both total cells and total libraries 
all_info_df |> 
  dplyr::count(final_annotation, name = "total_cells") |>
  dplyr::left_join(total_libraries, by = "final_annotation") |> 
  dplyr::arrange(desc(total_cells))
```

@@

@@

# get individual gene counts for all marker genes 
gene_cts <- logcounts(sce[genes, ]) |>
  as.matrix() |>
  t() |> 
  as.data.frame()
gene_cts$barcodes <- rownames(gene_cts)

@@

>>>>

- Annotations obtained by running `SingleR` with tumor cells as a reference output by [`aucell-singler-annotation.sh`](https://github.com/AlexsLemonade/OpenScPCA-analysis/blob/main/analyses/cell-type-ewings/aucell-singler-annotation.sh). 
- Consensus cell type annotations output by running [`assign-consensus-celltypes.sh` in `cell-type-consensus`](https://github.com/AlexsLemonade/OpenScPCA-analysis/blob/main/analyses/cell-type-consensus/assign-consensus-celltypes.sh). 
- AUC values as calculated by `AUCell` for a set of Ewing sarcoma specific gene sets from MSigDB output by [`run-aucell-ews-signatures.sh`](https://github.com/AlexsLemonade/OpenScPCA-analysis/blob/main/analyses/cell-type-ewings/run-aucell-ews-signatures.sh). 
- Custom gene sets from [`visser-all-marker-genes.tsv`](https://github.com/AlexsLemonade/OpenScPCA-analysis/blob/main/analyses/cell-type-ewings/references/visser-all-marker-genes.tsv), [`tumor-cell-state-markers.tsv`](https://github.com/AlexsLemonade/OpenScPCA-analysis/blob/main/analyses/cell-type-ewings/references/tumor-cell-state-markers.tsv), and [`combined-validation-markers.tsv`](https://github.com/AlexsLemonade/OpenScPCA-analysis/blob/main/analyses/cell-type-ewings/references/combined-validation-markers.tsv).

@@

Any cells that are unable to be labeled via consensus cell types are labeled as "Unknown" and I expect these will line up with cells labeled as tumor cells by `SingleR`.  
Note that only cell types from `SingleR` with > 50 cells are shown in the heatmap.

@@

```{r}
# get columns with mean expression of custom gene sets 
mean_exp_columns <- colnames(all_info_df) |> 
  stringr::str_subset("_mean$")

# make sure order is defined by cell type order
all_info_df <- all_info_df |> 
  dplyr::arrange(final_lumped)

@@

# first get tally for how many libraries each cell type is present in 
total_libraries <- all_info_df |> 
  dplyr::select(final_annotation, library_id) |> 
  unique() |> # get rid of duplicate cells
  dplyr::add_count(final_annotation, name = "total_libraries") |>
  dplyr::select(final_annotation, total_libraries)

# print out both total cells and total libraries 
all_info_df |> 
  dplyr::count(final_annotation, name = "total_cells") |>
  dplyr::left_join(total_libraries, by = "final_annotation") |> 
  dplyr::arrange(desc(total_cells)) |>
  unique()
```

@@

Note that only genes with mean expression > 0 and expressed in at least 10% of cells are included in the dotplot. 
All genes are included in the heatmap.

@@

# get individual gene counts for all marker genes 
gene_cts <- logcounts(sce[genes, ]) |>
  t() |> 
  as.data.frame()
gene_cts$barcodes <- rownames(gene_cts)

@@

The mean expression in this plot is capped at 2.5 for coloring purposes, but some values may be higher.
